Group Signatures Meet Anonymous Credentials: Or, How to Build Privacy Preserving Identities with Accountability

### 2022-02-27 prism-backend/research/gsac/uas-model.tex
First potentially stable model for UAS. I'm sure it will still change, though.

% Our notion of traceability for \UAS is inspired by both. First, we restrict to
% signatures that are accepted by \Verify. For any such signature, \Inspect has to
% create a valid output (i.e., must not abort). The output of \Inspect has to be
% accepted by \Judge too. Moreover, there must exist one user (either honest or
% corrupt) who used his public key to obtain one or more credentials that allowed
% him to produce a signature which, once inspected, returns \y.
In group signatures, traceability ensures that, in the presence of an honest
issuer, every signature accepted by the verify algorithm must have been created
by a user that joined the group.  With conventional opening, this is essentially
checked by requesting the adversary to produce a signature, opening it, and
checking whether or not it comes from a group member; i.e., this ensures that
inspection cannot be forged in the presence of an honest issuer.
%
The somehow equivalent property in anonymous credentials is unforgeability. It
requires that, again in the presence of an honest issuer, no adversary can
succeed in a credential presentation for attributes (or predicates of them) that
are not contained in a legitimately issued credential it controls. In other
words, that the evaluation of the credential attributes was not forged.
%
Both notions have an equivalent in \UAS. That is, we need to ensure that, in the
presence of an honest issuer, no signature accepted by \Verify can result in a
wrong inspection, nor it can have been obtained by a wrong evaluation of some
credential. Note, however, that \UAS offer the adversary yet another way to
break traceability/unforgeability. Namely, an adversary can attempt to
illegitimately obtain a credential that later enables it to produce untraceable
signatures; i.e., signatures that evaluate correctly, or for which \Inspect
returns the appropiate value, yet the adversary should not have been able to
obtain a credential that allowed it to produce such signatures. Our notion of
traceability/unforgeability thus challenges the adversary to produce a signature
that breaks any of those conditions. The formal
definition of the traceability experiment is given in \figref{fig:exp-uas-trace},
where $\Otrace \gets \lbrace\HU,\CU\rbrace\GEN,\IGEN,\OGEN,\OCORR,\OBTISS,
\ISSUE,\SIGN,\INSPECT$.

@@

\paragraph{Discussion on the generality of traceability in \UAS schemes.} %
\todo{Comment why is this more general than AC and GS independently.}

Note also that we have implicitly assumed an unforgeability relaxation in the
sense that we do not consider a forgery the fact that an adversary can obtain
credentials that do not meet the \fissue predicate \emph{as long as they are
  not used to produced untraceable signatures}. Indeed, this is in line with
previous work \cite[Section 3.3.3]{ckl+15}, and it does make sense as, in the
real world, being able to fraudulently obtain credentials that cannot be used
later to authenticate, does not pose a risk.

% Note that this notion of traceability is, again, strictly more general than
% both group signatures and anonymous credentials.
% %
% On the group signatures side,
% it all depends on the \Inspect function. If we set it to the traditional
% inspection function, where it returns the identity (index) of the signer, then
% our notion of traceability becomes roughly that of conventional group
% signatures\footnote{With the exception that, in \UAS, the same user can own
%   multiple credentials -- all of which must return the same identity}.
% On the other hand, \Inspect may be set to a non-injective function,
% meaning that it can be possible that different user public keys, credentials and
% signature tuples, may produce the same output. In that case, though, our notion
% of traceability ensures that there must exist \emph{at least one} user capable
% of issuing that signature. Indeed, this is complemented with the non-frameability
% property, that ensures that the adversary cannot produce a signature that
% inspects to a value that no corrupt user can produce.
% %

% Finally, note that we can only ensure consistency of the value returned by
% \finsp, for all users, as long as the issuer is honest. This follows from the
% fact that \finsp depends on the user key pair and the credentials that are
% issued to each user. Therefore, a dishonest issuer could create untraceable
% credentials (i.e., issue credentials without updating the \trans table), so that
% the \finsp function outputs whatever desired value, yet it cannot be traced to
% any registered user. That is, this minimality argument is just a generalization
% of the one described in \cite[Section 4.3]{bsz05}. This is also consistent with
% the unforgeability requirement of anonymous credentials systems, as a corrupt
% issuer can grant arbitrary credentials to any user, and therefore honest
% issuers are required.

\paragraph{Non-frameability.} %
Non-frameability variants are a core unforgeability-type property in group
signatures. However, no
similar property is modeled for anonymous credentials (\todo{see the discussion
  in \secref{sec:introduction} for further detail}). It is a quite strong
property, as it must be ensured even in the presence of dishonest issuer and
opener. Intuitively, it prevents the adversary from creating a signature that
frames an honest user. Depending on the inspection capabilities of the scheme,
this framing could be done in different ways; i.e., by convincing third parties
that signatures by different (possibly corrupt) users are linked, or directly
by having open proofs output the identity of a user who did not create the
signature being opened.
%

The notion of non-frameability in \UAS schemes is unavoidably more subtle than
in group signatures, though. To see this, we note that, by allowing any
arbitrary inspection function \finsp to be used, it can be perfectly valid to
have a signature produced by a corrupted user, output the same \y value upon
inspection than the one that \finsp outputs when inspecting a signature by an
honest user. As a concrete example, imagine an inspection function that returns
the banking account number (or cryptocurrency address) to be used to request
payment of a fine. In this example, a corrupted user and an honest user could
both have delegated this to some agency\footnote{\todo{Maybe, come up with a
    better example.}}, so it makes perfect sense that the inspection function
returns the same value for both users.
%
Thus, in \UAS schemes, we must be more subtle, by necessity.

In \UAS schemes, in order for a user to be framed \todo{...}. The
non-frameability property for \UAS schemes is formally defined in
\figref{fig:exp-uas-frame}, where $\Oframe \gets \lbrace\HU,\CU\rbrace\GEN,
\lbrace\II,\OO\rbrace\GEN,\lbrace\II,\OO\rbrace\CORR,\OBTAIN,\SIGN$

@@

\paragraph{Discussion.} %
\todo{Argue that our properties cover (and augment) a direct combination
  (even if possible) of their GS and AC counterparts. Specifically, anonymity
  of \GSAC is stronger than anonymity in GS and AC, and the same for trace and
  non-frame.}

@@

As an exercise, recall the Sign-Encrypt-Prove approach of group signatures in
\cite{bsz05}. Roughly, there, a users' membership credential is a signature by
the issuer of the user's public key. Then, a group signature is a signature
over the message to be group-signed, signed with the user's private key, an
encryption of the user's membership credential, and a NIZK proving that the
private key used to sign the message corresponds to the public key in the
encrypted membership credential. The opener just keeps a list of the public
keys and the ``identity'' (index within the member list) of the user. Then,
inspection (opening) simply consists of having the inspector decrypt the
encrypted membership credential in the group signature, and find the matching
public key in the members' list. Take the trivial inspection function
$\finsp(\upk,\cred,\msg,\sig) = \cred$. Inspection in an \UAS scheme that uses
this trivial \finsp function is equivalent to conventional opening. Then,
following this exercise, our non-frameability definition for this inspection
function would read as \todo{...}

\subsection{Transformation to Interactive Presentations}
\label{ssec:interactivetransform}

@@

>>>>

% Our notion of traceability for \UAS is inspired by both. First, we restrict to
% signatures that are accepted by \Verify. For any such signature, \Inspect has to
% create a valid output (i.e., must not abort). The output of \Inspect has to be
% accepted by \Judge too. Moreover, there must exist one user (either honest or
% corrupt) who used his public key to obtain one or more credentials that allowed
% him to produce a signature which, once inspected, returns \y.
In group signatures, traceability ensures that, in the presence of an honest
issuer, every signature accepted by the verify algorithm must have been created
by a user who joined the group.  With conventional opening, this is essentially
checked by requesting the adversary to produce a signature, opening it, and
checking whether or not it comes from a group member; i.e., this ensures that,
as long as there are verifiable inspections, inspection cannot be forged in the
presence of an honest issuer.
%
The somehow equivalent property in anonymous credentials is unforgeability. It
requires that, again in the presence of an honest issuer, no adversary can
succeed in a credential presentation for attributes (or predicates of them) that
are not contained in a legitimately issued credential it controls. In other
words, that the evaluation of the credential attributes cannot be forged.
%
Both notions have an equivalent in \UAS. That is, we need to ensure that, in the
presence of an honest issuer, no signature accepted by \Verify can result in a
wrong inspection, nor it can have been obtained by a wrong evaluation of some
credential. Note, however, that \UAS offer the adversary yet another way to
break traceability/unforgeability. Namely, an adversary can attempt to
illegitimately obtain a credential that later enables it to produce untraceable
signatures; i.e., signatures that evaluate correctly, or for which \Inspect
returns the appropiate value, yet the adversary should not have been able to
obtain a credential that allowed it to produce such signatures. Our notion of
traceability/unforgeability thus challenges the adversary to produce a signature
that breaks any of those conditions. The formal
definition of the traceability experiment is given in \figref{fig:exp-uas-trace},
where $\Otrace \gets \lbrace\HU,\CU\rbrace\GEN,\IGEN,\OGEN,\OCORR,\OBTISS,
\ISSUE,\SIGN,\INSPECT$.

@@

\paragraph{Discussion on the generality of traceability in \UAS schemes.} %
The notion of traceability we present for \UAS is strictly more general than
the corresponding one for group signatures. This is again the case as a direct
consequence of the fact that \Inspect, which can return an arbitrary function
of the signer's credential (and signed message), is a generalization of the
conventional \Open -- if we make \Inspect return just the identity of the
signer, then we get something similar to the usual notion of traceability.
Although, even in that case, we need to take into account attributes, and the
fact that the same user may obtain multiple credentials (that is why, even when
having \Inspect return the identity of the signer, our notion is not exactly
the same). And, in this sense, the unforgeability notion that \UAS have is
equivalent to that of anonymous credentials. It would seem, though, that we do
not need the traceability part of group signature; after all, it is the
protection against wrong claims on attributes what enables meaningful and
flexible authentication. However, the type of protection against misuses of
the inspection functionality that we can get with an honest issuer (as in
traceability) is much higher than without an honest user (as in
non-frameability). Specifically, with an honest issuer we can ensure that
the adversary cannot even alter the value returned by \Inspect on signatures
by corrupt users. Whereas, with a corrupt issuer, all we can ensure is that
the adversary cannot forge a signature from an honest user for which \Inspect
returns the same value as a signature by that honest user would produce.
Traceability is, therefore, a core property to ensure accountability
\todo{Discuss this in more detail somewhere else.}

Note also that we have implicitly assumed an unforgeability relaxation in the
sense that we do not consider a forgery the fact that an adversary can obtain
credentials that do not meet the \fissue predicate \emph{as long as they are
  not used to produced untraceable signatures}. Indeed, this is in line with
previous work \cite[Section 3.3.3]{ckl+15}, and it does make sense as, in the
real world, being able to fraudulently obtain credentials that cannot be later
used to authenticate, does not pose a risk. However, an adversary breaking
the issue policy defined by \fissue to get a credential and later using it to
create a valid signature (i.e., meeting \feval or \finsp) is considered a
forgery by our definition.

\paragraph{Non-frameability.} %
Non-frameability variants are a core unforgeability-type property in group
signatures. However, no
similar property is modeled for anonymous credentials (\todo{see the discussion
  in \secref{sec:introduction} for further detail}). It is a quite strong
property, as it must be ensured even in the presence of dishonest issuer and
opener. Intuitively, it prevents the adversary from creating a signature that
frames an honest user. Depending on the inspection capabilities of the scheme,
this framing could be done in different ways; i.e., by convincing third parties
that signatures by different (possibly corrupt) users are linked, or directly
by having inspection proofs output the identity of a user who did not create the
signature being inspected.

The notion of non-frameability in \UAS schemes is unavoidably more subtle than
in group signatures, though. To see this, we note that, by allowing any
arbitrary inspection function \finsp to be used, it can be perfectly valid to
have a signature produced by a corrupted user, output the same \y value upon
inspection than the one that \finsp outputs when inspecting a signature by an
honest user. As a concrete example, imagine an inspection function that returns
the banking account number (or cryptocurrency address) to be used to request
payment of a fine. In this example, a corrupted user and an honest user could
both have delegated this to some agency\footnote{\todo{Maybe, come up with a
    better example. How about spam filters that apply some sort of classification
    engine? \finsp could be this engine, and emails by many different users could
    be classified just as ``Spam'' or ``Not spam''}}, so it makes perfect sense
that the inspection function returns the same value for both users.
%
Thus, in \UAS schemes, we must be more subtle, by necessity. Note that we cannot
just require that no adversary can produce a signature for which inspection
returns the same value as it would for a signature honestly generated by an
uncorrupted user. This is again the case since inspection may be non-injective,
which means that inspection of a signature by some corrupted user may
legitimately produce the same value as that of an honest user. Thus, we need
to further refine the definition by requiring that (i) no adversary can produce
a signature that is recognized (via \Identify) as originating from an honest
user \uid, (ii) who owns a credential that makes \finsp return a valid \y value
(i.e., generated by \Inspect along with a proof, and accepted by \Judge), (iii)
without having queried the \SIGN oracle before. This non-frameability property
for \UAS schemes is formally defined in \figref{fig:exp-uas-frame}, where
$\Oframe \gets \lbrace\HU,\CU\rbrace\GEN,
\lbrace\II,\OO\rbrace\GEN,\lbrace\II,\OO\rbrace\CORR,\OBTAIN,\SIGN$

@@

\paragraph{Discussion on the generality of non-frameability in \UAS schemes.} %
Anonymous credentials do not have non-frameability property and, thus, it is
hard to make a comparison. However, we can draw some connections with AC schemes
that support revocation, as revocation is somehow equivalent to linking, which
is a type of inspection available in group signatures. In this sense, note that
basic revocation (without straight deanonymization) can be trivially achieved
through our generic \Inspect function. For instance, one could set \finsp to
be a pseudorandom number seeded with the user's public key (or credential). In
this sense, \Inspect could be essentially seen as a Verifiable Random Function.
If we compare with group signatures, our notion is again more general than the
conventional one. Indeed, in typical group signatures, \Open returns the
identity of the signer, and thus does non-frameability not allow the same value
to be returned when opening signatures by different users. Otherwise, a
corrupted user controlled by the adversary could be able to create a signature
opening to some honest user (hence, framing him). But this is again possible
with our definition which, in case of making \Inspect equal \Open, becomes
equivalent to conventional non-frameability. However, it also allows more
generic situations in which signatures by different users may return the same
value.

\subsection{Functionality ``Transformations''}
\label{ssec:transformations}

@@

\subsubsection{Transformation to Interactive Presentations}
\label{sssec:interactivetransform}

@@

\subsubsection{Combining Multiple Credentials in One Presentation}

\todo{I think this might be possible if we can use \fissue so that any user can
  create arbitrary groups, where he is the issuer, an can thus run
  \Obtain,\Issue protocols to combine credentials from other groups. This is
  probably inefficient (or, at least, more than doing so wihtout having to
  creat ad hoc groups), but at least allows the enhanced functionality...}
